---
title: "Mapping our cybersecurity AAS to NICE work roles"
subtitle: "A community-college program coordinator preparing for the CAE designation pathway"
description: "A program coordinator uses cybedtools to map cybersecurity AAS course outcomes to NICE work-role tasks for the National Centers of Academic Excellence designation."
categories: [Higher Ed, Workforce, NICE × CSEC2017]
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(here)
library(stringr)

best  <- readRDS(here("_data", "nice_csec2017_best.rds")) |>
  mutate(pct = round(similarity * 100, 1))
top3  <- readRDS(here("_data", "nice_csec2017_alignment.rds"))
```

## Marcus's situation

Marcus runs the cybersecurity AAS at a community college near a DoD installation. His program holds CAE-CD designation, granted in 2018, with re-designation work on the horizon. Marcus carries the documentation load.

His program's course outcomes were originally mapped to NIST's NICE Framework work-role tasks. NICE was the default cyber-workforce reference when the program was built. The articulation agreement with the regional state university's BS-Cybersecurity program references CSEC2017 Knowledge Areas instead. That program adopted CSEC2017 as its internal curriculum-mapping framework when it pursued ABET cybersecurity accreditation. When Marcus's curriculum committee asked whether his NICE-aligned AAS course outcomes implicitly cover the CSEC2017 KAs the receiving institution maps against, he pulled cybedtools to sanity-check the question before committing to a manual crosswalk.

::: {.column-margin}
This page uses real data computed against cybedtools v0.2.0. The
persona is composite. The question, the code, and the result are not.
:::

## The question, formally stated

For each of the 41 NICE work roles, what's the closest CSEC2017 Knowledge Area by full-document text similarity?

::: {.column-margin}
**Vocabulary overlap** is the percentage of unique words two units share, after dropping common words ("the," "and," "of") and short tokens. 0% means no shared vocabulary, 100% means identical wording. The underlying metric and the methodology choices live on the [analytic query page](../queries/nice-csec2017-alignment.qmd).
:::

## What cybedtools surfaces

```{r}
#| label: tbl-similarity-summary
#| tbl-cap: "Best-match vocabulary overlap across all 41 NICE work roles"
tibble::tibble(
  Statistic = c("Minimum", "Median", "Mean", "Maximum"),
  `Vocabulary overlap` = c(
    paste0(min(best$pct), "%"),
    paste0(median(best$pct), "%"),
    paste0(round(mean(best$pct), 1), "%"),
    paste0(max(best$pct), "%")
  )
) |>
  knitr::kable()
```

```{r}
#| label: tbl-ka-frequency
#| tbl-cap: "How often each CSEC2017 KA appears as some NICE role's top match"
best |>
  count(`CSEC2017 KA` = csec_name, name = "NICE roles top-matched") |>
  arrange(desc(`NICE roles top-matched`)) |>
  knitr::kable()
```

The first thing the data tells Marcus: every best-match overlap is below 10%. The strongest pairing in the entire set lands at roughly 7% vocabulary overlap. The median is 4%. Across all 41 NICE work roles, none has a CSEC2017 Knowledge Area whose vocabulary substantially matches its own.

The second thing: 27 of 41 NICE roles are pulled toward the same KA, Component Security. Three of the eight KAs (Data Security, Human Security, Societal Security) never appear as any NICE role's best match.

## What this means for Marcus

The crosswalk he was about to build does not exist at the vocabulary level. NICE describes positions, the work a person in this role performs. CSEC2017 describes curricular knowledge areas, the body of material a cyber program teaches. They speak different languages on purpose, and the data reflects that.

The dominance of Component Security as top match is a tell, not a finding. Component Security has the most general-purpose vocabulary of the eight KAs. Anything about hardware, firmware, integrated systems, or technology components falls in scope, so it sweeps up vocabulary-similar NICE roles for a structural reason, not a substantive one. Marcus should not read "27 NICE roles map to Component Security" as a curriculum-coverage claim.

### The committee gets a defensible answer

When the curriculum committee asks whether NICE-to-CSEC2017 alignment was investigated, Marcus can point at this page. The evidence-of-absence result moves the committee conversation to the right level: course-outcome to course-outcome equivalence with the receiving BS program, not framework-to-framework alignment.

### What Marcus should do instead

For the CAE-CD re-designation file, the formally-recognized alignment is between his course outcomes and the CAE Knowledge Units (the 2024 revision). The CSEC2017 reference in the CAE program criteria is thought-model context, not an alignment-proof requirement. cybedtools does not currently include CAE KUs. The KUs are governed by the NSA-led CAE program with CISA as co-sponsor and are not openly licensed for redistribution. The KU mapping is a separate workflow Marcus runs through the program criteria directly.

For the articulation defense, the move is course-outcome-to-course-outcome equivalence with the receiving institution, not framework-to-framework crosswalking. cybedtools' job here was to rule out the framework-to-framework path so Marcus knows where to spend his time.

### A note on DoD-side credentialing

If Marcus's program is also pursuing DoD-aligned credentialing pathways (CySP scholarships, NICCS-listed courses, defense-contractor MOUs), NICE-to-DCWF is a separate cybedtools query. It surfaces where DCWF v5.1's defense-side work roles share element identifiers with NICE's general-purpose work roles. The shared identifiers reflect genuine overlap at the task, knowledge, and skill level, but they do not imply role-equivalence. DCWF is owned by DoD CIO and adds defense-mission work-role context NICE roles do not carry. That query is for student credentialing-pathway advising, not for AAS program articulation.

## See also

- [The analytic query behind this scenario](../queries/nice-csec2017-alignment.qmd) walks through the SPARQL, the tokenization, and the similarity-metric choice.
- [NICE Framework page](../frameworks/nice-v2.qmd) for the framework's structure and licensing.
- [CSEC2017 page](../frameworks/csec2017.qmd) for the curricular-guidelines context.
- [The two-tier vocabulary](../concepts/vocabulary.qmd) for why `cybed:Role` (NICE work roles, ECSF profiles) and `cybed:OrganizingUnit` (CSEC2017 KAs and other non-role units) are distinct types.
- [K-12 alignment query](../queries/k12-alignment.qmd) for the analogous question at the secondary layer, useful if the AAS program has dual-enrollment HS feeders.
