---
title: "CyBOK v1.1.0"
subtitle: "Cyber Security Body of Knowledge"
jurisdiction: "UK"
framework-type: "Pedagogy"
framework-slug: "cybok-v1.1.0"
license: "OGL v3.0"
date: 2026-09-21
categories: [pedagogy, uk]
---

## At a glance

| | |
|---|---|
| Steward | National Cyber Security Centre (NCSC) |
| Canonical page | [Cyber Security Body of Knowledge](https://www.cybok.org/) |
| Version | 1.1.0 |
| Released | 2021-07 |
| Source format | PDF only (Introduction, per-Knowledge-Area trees, and an alphabetical index) |
| License | Crown Copyright, The National Cyber Security Centre 2021, licensed under the Open Government Licence v3.0 |
| `framework_summary` slug | `cybok-v1.1.0` |

## Structure in cybed: terms

::: {.column-page}

```{=html}
<figure class="schema-diagram">
<button type="button" class="toggle-all" aria-pressed="false" aria-label="Toggle schema names for all elements"><span class="toggle-icon" aria-hidden="true">▸</span><span class="toggle-text">schema names</span></button>
<svg viewBox="0 0 460 300" xmlns="http://www.w3.org/2000/svg" role="img" aria-labelledby="cybok-structure-title">
  <title id="cybok-structure-title">CyBOK structural mapping</title>
  <defs><marker id="cybok-arrow" viewBox="0 0 10 10" refX="8.5" refY="5" markerWidth="6" markerHeight="6" orient="auto-start-reverse"><path d="M0,0 L10,5 L0,10 z" fill="currentColor" opacity="0.55"/></marker></defs>
  <g stroke="currentColor" stroke-width="1.4" fill="none" opacity="0.55">
    <line x1="230" y1="58" x2="230" y2="100" marker-end="url(#cybok-arrow)"/>
    <line x1="230" y1="170" x2="230" y2="212" marker-end="url(#cybok-arrow)"/>
  </g>

  <g class="cyb-node" aria-label="Framework">
    <rect x="150" y="20" width="160" height="40" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="230" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Framework</text>
    <text class="lt" x="230" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">cybok:Framework</text>
  </g>

  <g class="cyb-node" aria-label="Knowledge Area">
    <rect x="70" y="100" width="320" height="70" rx="6" fill="rgba(14,165,233,0.16)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="230" y="130" text-anchor="middle" font-family="Inter,sans-serif" font-size="14" font-weight="600" fill="currentColor">Knowledge Area</text>
    <text class="lp" x="230" y="151" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.65">(21 KAs across 5 categories)</text>
    <text class="lt" x="230" y="125" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">cybok:KnowledgeArea</text>
    <text class="lt" x="230" y="145" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:OrganizingUnit</text>
    <text class="lt" x="230" y="161" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.55">(no cybed:Role)</text>
  </g>

  <g class="cyb-node" aria-label="Topic / Indicative Material">
    <rect x="90" y="212" width="280" height="60" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="230" y="240" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Topic / Indicative Material</text>
    <text class="lt" x="230" y="257" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">cybok:Topic, cybok:IndicativeMaterial = cybed:RoleElement</text>
  </g>

  <g class="cyb-edge-label" aria-label="Framework-to-KA connection">
    <text class="ep" x="238" y="84" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">contains</text>
    <text class="et" x="238" y="84" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasOrganizingUnit</text>
  </g>
  <g class="cyb-edge-label" aria-label="KA-to-topic connection">
    <text class="ep" x="238" y="196" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">has element</text>
    <text class="et" x="238" y="196" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasElement</text>
  </g>
</svg>
<figcaption>Topics come one layer below the Knowledge Tree's root; Indicative Material comes one layer further down, from a tree structure read off each Knowledge Tree diagram rather than from tabular data. Click any box or edge label to flip between plain English and the technical schema name.</figcaption>
</figure>
```

:::

## Counts

```{r}
#| label: setup
#| include: false
n <- cybedtools::framework_summary
n <- as.list(n[n$framework_slug == "cybok-v1.1.0", ])
cm <- function(x) format(x, big.mark = ",")
d1 <- function(x) format(round(x, 1), nsmall = 1)
```

```{r}
#| label: tbl-counts
tibble::tibble(
  Measure = c(
    "Knowledge Areas",
    "Strict elements (Topics)",
    "Subpoints (Indicative Material, parsed from Knowledge Trees)",
    "With-examples elements",
    "Elements per Knowledge Area, strict",
    "Elements per Knowledge Area, with-examples"
  ),
  Value = c(
    cm(n$organizing_unit_count),
    cm(n$element_count_strict),
    cm(n$subpoint_count),
    cm(n$element_count_with_examples),
    d1(n$elements_per_organizing_unit_strict),
    d1(n$elements_per_organizing_unit_with_examples)
  )
) |>
  knitr::kable()
```

CyBOK publishes no CSV, JSON, or XML: every count here comes from
reading a Knowledge Tree as a drawing. Node labels come from the PDF
text layer; parent-child links come from the drawing itself, recovered
by blanking out each node box and tracing the remaining ink. The gap
between the strict and with-examples counts reflects how much of
CyBOK's specification depth sits in Indicative Material terms one
layer below the Topic level, rather than in the Topics themselves.

## Provenance

::: {.provenance-cards .column-page}

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-archive"/></svg>

**Source**

The Introduction to CyBOK Knowledge Area (for KA names and categories), one Knowledge Tree PDF per KA (for Topics and Indicative Material), and the alphabetical Indicative Material index (for KA acronyms and cross-checking).
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-gear"/></svg>

**Ingestion**

`scripts/010-ingest-cybok.R` renders each Knowledge Tree page, blanks out the node boxes, and reconstructs the tree from the remaining connecting ink plus the PDF's own text-layer word coordinates for labels.
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-license"/></svg>

**License**

Crown Copyright, The National Cyber Security Centre 2021, licensed under the Open Government Licence v3.0, which permits copying, publishing, adapting, and commercial use on condition of attribution. Every source document (the Introduction, each Knowledge Tree, and the A-to-Z index) carries the same notice.
:::

:::

## Caveats

- Knowledge Tree parsing is **drawing recovery, not text extraction**. A KA whose tree uses an unusual layout is the most likely source of a structural error; cross-checking against the alphabetical index catches most but not all of these.
- CyBOK defines its own vocabulary for tree depth (**Topic**, then **Indicative Material** one layer further down); deeper layers in a given tree, where they exist, are not separately typed.
- Case is **preserved exactly as printed** in each source: Knowledge Trees mostly print Topics in lower case, and the A-to-Z index prints everything in capitals. Matching a Topic name across the two requires case-insensitive comparison.

## Related

- [Cross-framework density spread](../queries/density-spread.qmd)
- [The vocabulary](../concepts/vocabulary.qmd)
