---
title: "CyQUAL 1.2.0"
subtitle: "Czech Republic National Cybersecurity Qualifications Framework"
jurisdiction: "CZ"
framework-type: "Workforce"
framework-slug: "cyqual-v1.2.0"
license: "Open data, attribution required"
date: 2026-09-21
categories: [workforce, cz]
---

## At a glance

| | |
|---|---|
| Steward | CyQUAL, developed at Masaryk University |
| Canonical page | [CyQUAL platform](https://platform.cyqual.cz/) |
| Version | 1.2.0 |
| Source format | Open-data JSON export |
| License | Published as open data. Written confirmation from the CyQUAL team permits a derived RDF/JSON-LD graph, on condition of attribution to CyQUAL and to Masaryk University |
| `framework_summary` slug | `cyqual-v1.2.0` |

## Structure in cybed: terms

::: {.column-page}

```{=html}
<figure class="schema-diagram">
<button type="button" class="toggle-all" aria-pressed="false" aria-label="Toggle schema names for all elements"><span class="toggle-icon" aria-hidden="true">▸</span><span class="toggle-text">schema names</span></button>
<svg viewBox="0 0 460 400" xmlns="http://www.w3.org/2000/svg" role="img" aria-labelledby="cyqual-structure-title">
  <title id="cyqual-structure-title">CyQUAL structural mapping</title>
  <defs><marker id="cyqual-arrow" viewBox="0 0 10 10" refX="8.5" refY="5" markerWidth="6" markerHeight="6" orient="auto-start-reverse"><path d="M0,0 L10,5 L0,10 z" fill="currentColor" opacity="0.55"/></marker></defs>
  <g stroke="currentColor" stroke-width="1.4" fill="none" opacity="0.55">
    <line x1="230" y1="58" x2="230" y2="100" marker-end="url(#cyqual-arrow)"/>
    <line x1="230" y1="170" x2="230" y2="212" marker-end="url(#cyqual-arrow)"/>
    <line x1="230" y1="282" x2="230" y2="324" marker-end="url(#cyqual-arrow)"/>
  </g>

  <g class="cyb-node" aria-label="Framework">
    <rect x="150" y="20" width="160" height="40" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="230" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Framework</text>
    <text class="lt" x="230" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">cyqual:Framework</text>
  </g>

  <g class="cyb-node" aria-label="Category / specialization area">
    <rect x="60" y="100" width="340" height="70" rx="6" fill="rgba(14,165,233,0.16)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="230" y="130" text-anchor="middle" font-family="Inter,sans-serif" font-size="14" font-weight="600" fill="currentColor">Category, specialization area</text>
    <text class="lp" x="230" y="151" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.65">(two grouping tiers above the role)</text>
    <text class="lt" x="230" y="145" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:OrganizingUnit</text>
  </g>

  <g class="cyb-node" aria-label="Work role">
    <rect x="120" y="212" width="220" height="70" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="230" y="247" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Work role</text>
    <text class="lt" x="230" y="237" text-anchor="middle" font-family="Inter,sans-serif" font-size="12" font-weight="600" fill="currentColor">cyqual:WorkRole</text>
    <text class="lt" x="230" y="256" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:Role = cybed:OrganizingUnit</text>
  </g>

  <g class="cyb-node" aria-label="Task, requirement, competency">
    <rect x="80" y="324" width="300" height="60" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="230" y="352" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Task, Requirement, Competency</text>
    <text class="lt" x="230" y="367" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:RoleElement</text>
  </g>

  <g class="cyb-edge-label" aria-label="Framework-to-category connection">
    <text class="ep" x="238" y="84" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">contains</text>
    <text class="et" x="238" y="84" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasOrganizingUnit</text>
  </g>
  <g class="cyb-edge-label" aria-label="Category-to-work-role connection">
    <text class="ep" x="238" y="196" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">contains</text>
    <text class="et" x="238" y="196" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasOrganizingUnit</text>
  </g>
  <g class="cyb-edge-label" aria-label="Work-role-to-element connection">
    <text class="ep" x="238" y="308" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">has element</text>
    <text class="et" x="238" y="308" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasElement</text>
  </g>
</svg>
  <figcaption>Category and specialization area nest above the work role; tasks and requirements (the latter classified by a separate competency/competency-group axis) attach below it. Click any box or edge label to flip between plain English and the technical schema name.</figcaption>
</figure>
```

:::

## Counts

```{r}
#| label: setup
#| include: false
n <- cybedtools::framework_summary
n <- as.list(n[n$framework_slug == "cyqual-v1.2.0", ])
cm <- function(x) format(x, big.mark = ",")
d1 <- function(x) format(round(x, 1), nsmall = 1)
```

```{r}
#| label: tbl-counts
tibble::tibble(
  Measure = c(
    "Organizing units (categories, specialization areas, work roles)",
    "Work roles",
    "Strict elements (tasks and requirements)",
    "Subpoints (parsed enumeration lists)",
    "With-examples elements",
    "Elements per organizing unit, strict",
    "Elements per work role, strict",
    "Elements per work role, with-examples"
  ),
  Value = c(
    cm(n$organizing_unit_count),
    cm(n$role_count),
    cm(n$element_count_strict),
    cm(n$subpoint_count),
    cm(n$element_count_with_examples),
    d1(n$elements_per_organizing_unit_strict),
    d1(n$elements_per_role_strict),
    d1(n$elements_per_role_with_examples)
  )
) |>
  knitr::kable()
```

The source hierarchy is category to specialization area to work role,
with tasks and requirements attached to work roles through two join
tables that carry no proficiency level, weight, or mandatory flag:
every association is an unqualified membership edge. Requirements are
additionally classified by a competency and competency-group axis
that sits alongside, not beneath, the work-role hierarchy.

## Provenance

::: {.provenance-cards .column-page}

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-archive"/></svg>

**Source**

CyQUAL's official open-data JSON export, version 1.2.0, retrieved 2026-09-16.
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-gear"/></svg>

**Ingestion**

`scripts/010-ingest-cyqual.R` flattens seven entity arrays and two association arrays into tidy tables and parses the structure into the cybed schema.
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-license"/></svg>

**License**

The open-data JSON export carries no rights field of any kind. The CyQUAL team at Masaryk University confirmed in writing on 2026-09-16 that the framework is published as open data and publicly available, that a derived RDF/JSON-LD graph may be built from it, and that the only condition is attribution to CyQUAL and to Masaryk University.
:::

:::

## Caveats

- CyQUAL is ingested from its **English-language export when available**; the Czech export is used otherwise. No translation happens in the pipeline itself. Provenance records which language a given ingestion run used.
- **Neither join table carries a strength-of-relationship signal.** A task or requirement attached to a work role is a membership edge only; nothing in the source distinguishes a core task from a peripheral one.
- Organizing-unit counts mix **three structurally different tiers** (category, specialization area, work role). Analyses comparing "organizing units" across frameworks should filter to `cyqual:WorkRole` for a role-level comparison.

## Related

- [Cross-framework density spread](../queries/density-spread.qmd)
- [The vocabulary](../concepts/vocabulary.qmd)
