---
title: "DCWF v5.1"
subtitle: "DoD Cyber Workforce Framework (DoDD 8140)"
jurisdiction: "US"
framework-type: "Workforce"
framework-slug: "dcwf-v5.1"
license: "Public domain"
date: 2026-05-08
categories: [workforce, us]
---

## At a glance

| | |
|---|---|
| Steward | DoD Chief Information Officer |
| Canonical page | [DoD Cyber Workforce Framework](https://dodcio.defense.gov/Cyber-Workforce/DCWF/) |
| Version | 5.1 |
| Released | 2025-07-25 |
| Aligned to | NICE Framework |
| Source format | DoD CIO Work Role Tool (XLSX) |
| License | A US Government work prepared by DoD personnel, so not subject to copyright in the United States under 17 U.S.C. 105. The v5.1 workbook carries no distribution statement |
| `framework_summary` slug | `dcwf-v5.1` |

## Structure in cybed: terms

::: {.column-page}

```{=html}
<figure class="schema-diagram">
<button type="button" class="toggle-all" aria-pressed="false" aria-label="Toggle schema names for all elements"><span class="toggle-icon" aria-hidden="true">▸</span><span class="toggle-text">schema names</span></button>
<svg viewBox="0 0 540 310" xmlns="http://www.w3.org/2000/svg" role="img" aria-labelledby="dcwf-structure-title">
  <title id="dcwf-structure-title">DCWF structural mapping</title>
  <defs><marker id="dcwf-arrow" viewBox="0 0 10 10" refX="8.5" refY="5" markerWidth="6" markerHeight="6" orient="auto-start-reverse"><path d="M0,0 L10,5 L0,10 z" fill="currentColor" opacity="0.55"/></marker></defs>
  <g stroke="currentColor" stroke-width="1.4" fill="none" opacity="0.55">
    <line x1="270" y1="58" x2="270" y2="100" marker-end="url(#dcwf-arrow)"/>
    <line x1="220" y1="172" x2="160" y2="220" marker-end="url(#dcwf-arrow)"/>
    <line x1="320" y1="172" x2="380" y2="220" marker-end="url(#dcwf-arrow)"/>
  </g>

  <g class="cyb-node" aria-label="Framework">
    <rect x="190" y="20" width="160" height="40" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="270" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Framework</text>
    <text class="lt" x="270" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">dcwf:Framework</text>
  </g>

  <g class="cyb-node" aria-label="Work role">
    <rect x="110" y="100" width="320" height="72" rx="6" fill="rgba(14,165,233,0.16)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="270" y="142" text-anchor="middle" font-family="Inter,sans-serif" font-size="14" font-weight="600" fill="currentColor">Work role</text>
    <text class="lt" x="270" y="124" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">dcwf:WorkRole</text>
    <text class="lt" x="270" y="143" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:Role</text>
    <text class="lt" x="270" y="159" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:OrganizingUnit</text>
  </g>

  <g class="cyb-node" aria-label="Task">
    <rect x="80" y="220" width="160" height="60" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="160" y="255" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Task</text>
    <text class="lt" x="160" y="245" text-anchor="middle" font-family="Inter,sans-serif" font-size="12" font-weight="600" fill="currentColor">dcwf:Task</text>
    <text class="lt" x="160" y="263" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:RoleElement</text>
  </g>

  <g class="cyb-node" aria-label="KSA">
    <rect x="300" y="220" width="160" height="60" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="380" y="255" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">KSA</text>
    <text class="lt" x="380" y="245" text-anchor="middle" font-family="Inter,sans-serif" font-size="12" font-weight="600" fill="currentColor">dcwf:KSA</text>
    <text class="lt" x="380" y="263" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:RoleElement</text>
  </g>

  <g class="cyb-edge-label" aria-label="Framework-to-work-role connection">
    <text class="ep" x="278" y="84" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">contains</text>
    <text class="et" x="278" y="84" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasOrganizingUnit</text>
  </g>
  <g class="cyb-edge-label" aria-label="Work-role-to-element connection">
    <text class="ep" x="262" y="200" text-anchor="end" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">has element</text>
    <text class="et" x="262" y="200" text-anchor="end" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasElement</text>
  </g>
</svg>
<figcaption>Click any box or edge label to flip between plain English and the technical schema name.</figcaption>
</figure>
```

:::

## Counts

```{r}
#| label: setup
#| include: false
n <- cybedtools::framework_summary
n <- as.list(n[n$framework_slug == "dcwf-v5.1", ])
u <- readRDS(here::here("_data", "unit_subtypes.rds"))$n_other_frameworks
cm <- function(x) format(x, big.mark = ",")
d1 <- function(x) format(round(x, 1), nsmall = 1)
```

```{r}
#| label: tbl-counts
tibble::tibble(
  Measure = c(
    "Work roles",
    "Strict elements",
    "Subpoints (parsed enumeration lists)",
    "With-examples elements",
    "Elements per work role, strict",
    "Elements per work role, with-examples"
  ),
  Value = c(
    cm(n$role_count),
    cm(n$element_count_strict),
    cm(n$subpoint_count),
    cm(n$element_count_with_examples),
    d1(n$elements_per_role_strict),
    d1(n$elements_per_role_with_examples)
  )
) |>
  knitr::kable()
```

DCWF has no `cybed:Example` nodes -- its numbered statements carry no
separate prose-embedded clarifications the way Cyber.org K-12 or CSTA
do. But `r cm(n$subpoint_count)` of its Task/KSA statements phrase part of their content
as an enumerated list ("including X and Y", "(e.g., X, Y, and Z)"),
and the uniform Subpoint parser splits those out the same way it does
across the other `r cm(u)` frameworks in the corpus, so strict and with-examples do not
coincide here. This is DCWF's second correction of the day: a parser
bug fixed 2026-08-14 (a dead `e.g.` introducer pattern, plus a
soft-wrap newline being read as a false sentence boundary) meant most
of this framework's `(e.g., ...)`-style enumerations were previously
invisible to the Subpoint layer entirely. DCWF now carries the highest
with-examples density of any framework in the corpus, ahead of NICE.

## Provenance

::: {.provenance-cards .column-page}

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-archive"/></svg>

**Source**

DoD Cyber Workforce Framework reference materials.
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-gear"/></svg>

**Ingestion**

`scripts/010-ingest-dcwf.R` parses the source structure into the cybed schema.
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-license"/></svg>

**License**

A work of the United States Government prepared by DoD personnel, so not subject to copyright in the United States under 17 U.S.C. 105. The v5.1 workbook carries no distribution statement, so nothing in the artifact either grants or restricts distribution. The package does not bundle the source workbook. Users stage it locally.
:::

:::

## Caveats

- DCWF aligns to NICE and **shares element identifiers across the two frameworks**. Summing element counts across DCWF and NICE will double-count shared identifiers. Filter on `framework_slug` when summing.
- The ingested data reflects the **source export retrieved at ingestion time**. Re-run the ingestion script against a newer export for subsequent versions.
- DCWF is **authoritative for the DoD cyberspace workforce, not the civilian side**. Civilian federal and private-sector hiring pipelines that look DCWF-shaped typically map back through NICE.

## Related

- [NICE Framework v2](nice-v2.qmd), the framework DCWF aligns to.
- [Cross-framework density spread](../queries/density-spread.qmd)
- [The vocabulary](../concepts/vocabulary.qmd)
