---
title: "Frameworks"
subtitle: "What each framework specifies, where it sits, and what its source license permits"
---

```{r}
#| label: setup
#| include: false
suppressPackageStartupMessages({
  library(dplyr)
  library(reactable)
})
fs <- cybedtools::framework_summary
n_fw <- nrow(fs)
n_juris <- length(unique(fs$jurisdiction))
n_us <- sum(fs$jurisdiction == "US")
n_ped <- sum(fs$framework_type == "pedagogy")
ped_juris <- paste(sort(unique(fs$jurisdiction[fs$framework_type == "pedagogy"])), collapse = ", ")

# The `Units`/`Elements` columns must come from framework_summary, computed
# at render time, never hand-typed per-page frontmatter -- a listing block
# reading YAML frontmatter cannot do that, so this table is built directly
# in R instead. `framework-slug` in each page's own frontmatter is the join
# key; the href map below is the one place the slug-to-filename
# correspondence has to be spelled out, since filenames don't all match
# their slug (`digcomp-30.qmd` for `digcomp-3.0`, `ccssf.qmd` for
# `ccssf-2022`, and so on).
page_of_slug <- c(
  "ccssf-2022"        = "ccssf.qmd",
  "csec2017-v1"       = "csec2017.qmd",
  "csta-2017"         = "csta-2017.qmd",
  "csta-2026"         = "csta-2026.qmd",
  "cyberorg-k12-v1.0" = "cyberorg-k12.qmd",
  "cybok-v1.1.0"      = "cybok.qmd",
  "cyqual-v1.2.0"     = "cyqual.qmd",
  "dcwf-v5.1"         = "dcwf-v51.qmd",
  "digcomp-3.0"       = "digcomp-30.qmd",
  "ecsf-v1"           = "ecsf-v1.qmd",
  "nice-v2"           = "nice-v2.qmd",
  "otccf-v1.1"        = "otccf.qmd",
  "scywf-1.5"         = "scywf.qmd",
  "sfia-9"            = "sfia-9.qmd"
)

fw_table <- fs |>
  mutate(
    Framework  = framework_name,
    href       = sub("\\.qmd$", ".html", page_of_slug[framework_slug]),
    Jurisdiction = jurisdiction,
    Type       = tools::toTitleCase(framework_type),
    Units      = organizing_unit_count,
    Elements   = element_count_with_examples,
    License    = license
  ) |>
  arrange(framework_type, jurisdiction, framework_name) |>
  select(Framework, href, Jurisdiction, Type, Units, Elements, License)
```

```{r}
#| label: tbl-frameworks
reactable(
  fw_table,
  columns = list(
    href = colDef(show = FALSE),
    Framework = colDef(cell = function(value, index) {
      htmltools::tags$a(href = fw_table$href[index], value)
    })
  ),
  sortable = TRUE,
  filterable = FALSE,
  striped = TRUE,
  highlight = TRUE,
  bordered = FALSE,
  compact = TRUE,
  defaultPageSize = 10
)
```

The framework corpus splits along two axes: workforce versus
pedagogy, and jurisdiction, which now takes `r n_juris` values
(`r paste(sort(unique(fs$jurisdiction)), collapse = ", ")`). The
combinations are uneven. The US carries `r n_us` of the `r n_fw`
frameworks, and the `r n_ped` pedagogy frameworks sit in only
`r length(unique(fs$jurisdiction[fs$framework_type == "pedagogy"]))`
jurisdictions (`r ped_juris`). Together the frameworks span
the practical comparison surface for most cybersecurity workforce
and education research questions.

::: {.column-margin}
Counts on this table use the with-examples figure. Strict counts
(numbered standards only) appear on each framework's page and on
the [density-spread query](../queries/density-spread.qmd).
:::
