---
title: "NICE Framework v2"
subtitle: "Workforce Framework for Cybersecurity (NIST SP 800-181r1)"
jurisdiction: "US"
framework-type: "Workforce"
organizing-units: 53
elements: 2225
license: "Public domain"
date: 2026-05-08
date-modified: 2026-05-08
categories: [workforce, us]
---

## At a glance

| | |
|---|---|
| Steward | NIST (National Initiative for Cybersecurity Careers and Studies) |
| Canonical page | [NICE Framework Resource Center](https://www.nist.gov/itl/applied-cybersecurity/nice/nice-framework-resource-center) |
| Version | 2.2.0 |
| Released | 2026-04-28 |
| Source format | NIST Cybersecurity and Privacy Reference Tool (CPRT) JSON export |
| License | A US Government work, not subject to copyright in the United States under 17 U.S.C. 105. NIST reserves foreign rights and grants them back royalty-free worldwide, including derivative works |
| `framework_summary` slug | `nice-v2` |

## Structure in cybed: terms

::: {.column-page}

```{=html}
<figure class="schema-diagram">
<button type="button" class="toggle-all" aria-pressed="false" aria-label="Toggle schema names for all elements"><span class="toggle-icon" aria-hidden="true">▸</span><span class="toggle-text">schema names</span></button>
<svg viewBox="0 0 680 310" xmlns="http://www.w3.org/2000/svg" role="img" aria-labelledby="nice-structure-title">
  <title id="nice-structure-title">NICE Framework structural mapping</title>
  <defs><marker id="nice-arrow" viewBox="0 0 10 10" refX="8.5" refY="5" markerWidth="6" markerHeight="6" orient="auto-start-reverse"><path d="M0,0 L10,5 L0,10 z" fill="currentColor" opacity="0.55"/></marker></defs>
  <g stroke="currentColor" stroke-width="1.4" fill="none" opacity="0.55">
    <line x1="340" y1="58" x2="340" y2="100" marker-end="url(#nice-arrow)"/>
    <line x1="220" y1="172" x2="130" y2="220" marker-end="url(#nice-arrow)"/>
    <line x1="340" y1="172" x2="340" y2="220" marker-end="url(#nice-arrow)"/>
    <line x1="460" y1="172" x2="550" y2="220" marker-end="url(#nice-arrow)"/>
  </g>

  <g class="cyb-node" aria-label="Framework">
    <rect x="260" y="20" width="160" height="40" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="340" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Framework</text>
    <text class="lt" x="340" y="45" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">nice:Framework</text>
  </g>

  <g class="cyb-node" aria-label="Work role">
    <rect x="180" y="100" width="320" height="72" rx="6" fill="rgba(14,165,233,0.16)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="340" y="142" text-anchor="middle" font-family="Inter,sans-serif" font-size="14" font-weight="600" fill="currentColor">Work role</text>
    <text class="lt" x="340" y="124" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">nice:WorkRole</text>
    <text class="lt" x="340" y="143" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:Role</text>
    <text class="lt" x="340" y="159" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:OrganizingUnit</text>
  </g>

  <g class="cyb-node" aria-label="Task">
    <rect x="50" y="220" width="160" height="60" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="130" y="255" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Task</text>
    <text class="lt" x="130" y="245" text-anchor="middle" font-family="Inter,sans-serif" font-size="12" font-weight="600" fill="currentColor">nice:Task</text>
    <text class="lt" x="130" y="263" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:RoleElement</text>
  </g>

  <g class="cyb-node" aria-label="Knowledge statement">
    <rect x="260" y="220" width="160" height="60" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="340" y="255" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Knowledge statement</text>
    <text class="lt" x="340" y="245" text-anchor="middle" font-family="Inter,sans-serif" font-size="11.5" font-weight="600" fill="currentColor">nice:KnowledgeStatement</text>
    <text class="lt" x="340" y="263" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:RoleElement</text>
  </g>

  <g class="cyb-node" aria-label="Skill statement">
    <rect x="470" y="220" width="160" height="60" rx="6" fill="rgba(14,165,233,0.10)" stroke="#0284C7" stroke-width="1.6"/>
    <text class="lp" x="550" y="255" text-anchor="middle" font-family="Inter,sans-serif" font-size="13" font-weight="600" fill="currentColor">Skill statement</text>
    <text class="lt" x="550" y="245" text-anchor="middle" font-family="Inter,sans-serif" font-size="12" font-weight="600" fill="currentColor">nice:SkillStatement</text>
    <text class="lt" x="550" y="263" text-anchor="middle" font-family="Inter,sans-serif" font-size="10.5" fill="currentColor" opacity="0.7">= cybed:RoleElement</text>
  </g>

  <g class="cyb-edge-label" aria-label="Framework-to-work-role connection">
    <text class="ep" x="348" y="84" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">contains</text>
    <text class="et" x="348" y="84" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasOrganizingUnit</text>
  </g>
  <g class="cyb-edge-label" aria-label="Work-role-to-element connection">
    <text class="ep" x="332" y="200" text-anchor="end" font-family="Inter,sans-serif" font-size="11.5" fill="currentColor" opacity="0.7">has element</text>
    <text class="et" x="332" y="200" text-anchor="end" font-family="JetBrains Mono,monospace" font-size="10" fill="currentColor" opacity="0.7">cybed:hasElement</text>
  </g>
</svg>
<figcaption>Click any box or edge label to flip between plain English and the technical schema name.</figcaption>
</figure>
```

:::

## Counts

```{r}
#| label: setup
#| include: false
n <- cybedtools::framework_summary
n <- as.list(n[n$framework_slug == "nice-v2", ])
cm <- function(x) format(x, big.mark = ",")
d1 <- function(x) format(round(x, 1), nsmall = 1)
```

```{r}
#| label: tbl-counts
tibble::tibble(
  Measure = c(
    "Organizing units (work roles plus competency areas)",
    "Work roles",
    "Competency areas (second grouping axis, v2.2.0)",
    "Strict elements (numbered statements)",
    "Subpoints (parsed enumeration lists)",
    "Examples (Clarification-statement scaffolding)",
    "With-examples elements",
    "Elements per organizing unit, strict",
    "Elements per organizing unit, with-examples",
    "Elements per work role, strict",
    "Elements per work role, with-examples"
  ),
  Value = c(
    cm(n$organizing_unit_count),
    cm(n$role_count),
    cm(n$organizing_unit_count - n$role_count),
    cm(n$element_count_strict),
    cm(n$subpoint_count),
    cm(n$example_count),
    cm(n$element_count_with_examples),
    d1(n$elements_per_organizing_unit_strict),
    d1(n$elements_per_organizing_unit_with_examples),
    d1(n$elements_per_role_strict),
    d1(n$elements_per_role_with_examples)
  )
) |>
  knitr::kable()
```

Two denominators sit in that table and they answer different questions. The cross-framework density chart uses organizing units, which for NICE v2.2.0 means the `r n$role_count` work roles plus the `r n$organizing_unit_count - n$role_count` competency areas. The per-role figures hold the competency areas out of both numerator and denominator. Competency areas are a second grouping axis over the same knowledge and skill catalog rather than an additional set of roles.

NICE's strict and with-examples counts are identical. The framework
encodes detail entirely in numbered statements, so the ingestion has
no prose-embedded clarifications or annex examples to lift into
separate `cybed:Example` nodes.

## Provenance

::: {.provenance-cards .column-page}

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-archive"/></svg>

**Source**

NIST's CPRT JSON export. Retrieval URL and reference checksum live in the package's `provenance.yml`.
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-gear"/></svg>

**Ingestion**

`scripts/010-ingest-nice.R` parses the CPRT structure into the cybed schema. Stage the JSON file at `data/raw/nice/cprt-export.json` per `docs/framework-data-sources.md`.
:::

::: {.provenance-card}
<svg class="provenance-icon" aria-hidden="true"><use href="#icon-license"/></svg>

**License**

A work of the United States Government, not subject to copyright in the United States under 17 U.S.C. 105. NIST reserves foreign rights and grants them back, giving the public a non-exclusive, perpetual, paid-up, royalty-free, worldwide right to reprint the work in all formats and in derivative works. Attribute NIST as the source and do not imply NIST endorsement. The package does not bundle the source JSON. Users stage it locally.
:::

:::

## Caveats

- DCWF aligns to NICE and **shares element identifiers across the two frameworks**. Summing element counts across NICE and DCWF will double-count shared identifiers. Filter on `framework_slug` when summing.
- The ingested data reflects the **CPRT export retrieved at ingestion time**. Re-run the ingestion script against a newer export to pick up subsequent NICE versions.
- A small number of work roles **carry a disproportionate share of element coverage**. See the [top NICE work roles query](../queries/top-work-roles.qmd).

## Related

- [DCWF v5.1](dcwf-v51.qmd), the DoD's NICE-aligned counterpart.
- [Cross-framework density spread](../queries/density-spread.qmd),
  where NICE sits in the corpus ranking.
- [Top NICE work roles](../queries/top-work-roles.qmd), the power-law
  distribution across the `r n$role_count` work roles.
- [The vocabulary](../concepts/vocabulary.qmd), on how `nice:WorkRole`
  relates to `cybed:Role` and `cybed:OrganizingUnit`.
