---
title: "How does element density vary across frameworks?"
subtitle: "Per-unit density, strict and with examples"
description: "Per-unit element density across the framework corpus, strict and with examples. The twelvefold-spread finding, reframed as a comparison of design philosophy rather than a quality ranking."
categories: [Cross-framework]
---

```{r}
#| label: setup
#| include: false
library(dplyr)
library(ggplot2)
library(forcats)
library(here)

data_path <- here("_data", "density_spread.rds")
if (file.exists(data_path)) {
  density_spread <- readRDS(data_path)
} else {
  # Fallback: pull straight from the lazy-loaded package tibble so the
  # site renders even if the data-prep script has not been run.
  density_spread <- cybedtools::framework_summary |>
    arrange(desc(elements_per_organizing_unit_with_examples)) |>
    select(
      framework_name,
      framework_slug,
      framework_type,
      jurisdiction,
      organizing_unit_count,
      element_count_strict,
      element_count_with_examples,
      elements_per_organizing_unit_strict,
      elements_per_organizing_unit_with_examples,
      license
    )
}
```

## The question

How many elements does each framework specify per top-level organizing unit, and how does that count change when prose-encoded examples are included alongside numbered standards?

## Why it matters

Per-unit density is easy to mis-read as a quality ranking. It isn't one. What the figure measures is how granular each framework's stewards chose to be when they decided what work, what knowledge, or what competence belongs to a single organizing unit. NICE specifies tasks, knowledge statements, and skill statements per work role at one level of granularity. DigComp specifies competence areas at a different level, designed for citizen self-assessment. Both are valid specifications answering different questions.

## The result

```{r}
#| label: tbl-density
#| tbl-cap: "Eight frameworks, ranked by elements per organizing unit (with examples)"
density_spread |>
  transmute(
    Framework = framework_name,
    Type = tools::toTitleCase(framework_type),
    Jurisdiction = jurisdiction,
    Units = organizing_unit_count,
    `Strict elements` = element_count_strict,
    `With examples` = element_count_with_examples,
    `Per unit (strict)` = round(elements_per_organizing_unit_strict, 1),
    `Per unit (with ex.)` = round(elements_per_organizing_unit_with_examples, 1)
  ) |>
  knitr::kable()
```

```{r}
#| label: fig-density
#| fig-cap: "Per-unit element density across the framework corpus. Bars show the with-examples figure. The vertical tick on each bar marks the strict count for the same framework."
#| fig-width: 7.5
#| fig-height: 4.8
density_spread |>
  mutate(
    framework_name = fct_reorder(
      framework_name,
      elements_per_organizing_unit_with_examples
    )
  ) |>
  ggplot(aes(
    x = framework_name,
    y = elements_per_organizing_unit_with_examples,
    fill = jurisdiction
  )) +
  geom_col(width = 0.7) +
  geom_point(
    aes(y = elements_per_organizing_unit_strict),
    shape = 124, size = 6, color = "#0F172A"
  ) +
  coord_flip() +
  scale_fill_manual(
    values = c(US = "#0284C7", EU = "#0EA5E9", global = "#7DD3FC")
  ) +
  labs(
    x = NULL,
    y = "Elements per organizing unit",
    fill = "Jurisdiction"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 9, face = "bold"),
    axis.text.y = element_text(size = 10),
    plot.margin = margin(8, 16, 8, 8)
  )
```

## What this tells us

Three observations carry weight.

### The spread is real and large

The headline figure is roughly twelvefold (NICE 51.6 against DigComp 4.2, examples included). The strict-count spread is forty-sevenfold (NICE 51.6 against Cyber.org K-12 1.1, examples excluded). Either way, the gap merits an explanation.

### The spread reflects design philosophy

NICE was built to be operationally specific. It supports workforce planning at the position-description level for the United States federal government, and the granularity is the point. DigComp was built for citizen self-assessment, so the unit of specification is necessarily coarser. SFIA sits in the middle on purpose, the legacy of its history as a global IT-skills catalog. Not as fine-grained as NICE. Not as coarse as DigComp.

### The strict-vs-with-examples gap is informative on its own

Frameworks that encode pedagogical detail in prose clarification statements (Cyber.org K-12, CSTA) widen substantially when those examples are folded in. Workforce frameworks (NICE, DCWF, ECSF, SFIA) don't move, because every bit of detail they ship is already in a numbered statement. Read the gap as a proxy for how much pedagogical scaffolding lives outside the numbered taxonomy.

## What this doesn't tell us

### Density is not quality

A framework with high per-unit density isn't better-specified than one with low per-unit density. The denominator (organizing units) is a design choice, not a corpus property, and different denominators rerank the list.

### Strict counts undercount frameworks that encode detail in prose

The 47x strict spread overstates the specificity gap because it ignores the prose-encoded detail in the pedagogy frameworks. Use the with-examples figure when the comparison crosses the workforce-pedagogy boundary.

### Cross-framework comparisons require shared denominators

"Top-level organizing unit per framework" is the denominator used here. A NICE work role and a DigComp competence area aren't the same kind of object, and treating them as commensurable for the purpose of a single ratio is itself an analytic move worth examining before leaning on the number.

## Reproduce this

```{.r}
library(cybedtools)
library(dplyr)

framework_summary |>
  arrange(desc(elements_per_organizing_unit_with_examples)) |>
  select(
    framework_name,
    organizing_unit_count,
    element_count_strict,
    element_count_with_examples,
    elements_per_organizing_unit_strict,
    elements_per_organizing_unit_with_examples
  )
```

The `framework_summary` tibble ships with the package as a lazy-loaded data object. It is built from the canonical pipeline run documented in `data-raw/build-framework-summary.R` and tagged at each release.

## Related queries

- [How is element coverage distributed across US and EU frameworks?](us-eu-asymmetry.qmd), the 14:1 ratio and what it reframes.
- [Which NICE work roles concentrate the most element coverage?](top-work-roles.qmd), power-law shape inside NICE.
- [How does the four-framework K-12 alignment compare?](k12-alignment.qmd), pedagogy lens.
- [The vocabulary](../concepts/vocabulary.qmd), why `cybed:OrganizingUnit` is the right denominator.
