---
title: "Quick start"
subtitle: "Install the package, stage framework data, run your first query"
---

```{r}
#| label: setup
#| include: false
fl <- cybedtools::framework_licenses
permission_slugs <- c("cyqual-v1.2.0", "ccssf-2022", "otccf-v1.1")
n_granted <- sum(fl$granted[fl$slug %in% permission_slugs])
```

## Install cybedtools

Install from GitHub via `remotes` or `pak`:

```{.r}
# Option A: remotes
install.packages("remotes")
remotes::install_github("ryanstraight/cybedtools")

# Option B: pak (faster on first install)
install.packages("pak")
pak::pkg_install("ryanstraight/cybedtools")
```

## Stage the framework data

`cybedtools` does not redistribute framework source text. Each
framework's license governs how its source data is obtained and
stored. `docs/framework-data-sources.md` documents the canonical
retrieval URL, license, and SHA256 reference checksum for each
framework in the corpus.

```{.r}
# After install, see where to put each framework's source data:
system.file("doc", "framework-data-sources.md", package = "cybedtools") |>
  readLines() |> head(40)
```

For NICE and DCWF (US Government works, not subject to US
copyright under 17 U.S.C. 105), the package
ships pointers to the canonical NIST CPRT and DoD CIO downloads. For
SFIA, ECSF, Cyber.org K-12, CSTA, CSEC2017, and DigComp, follow the
steward's redistribution policy as documented per framework. CyQUAL,
CCSSF, and OTCCF are in the corpus on their stewards' terms,
`r n_granted` by written permission and CCSSF referenced as its
steward asked, and each carries its own terms. Read the per-framework
entries before staging those three.

## Run the pipeline

Once source data is staged in `data/raw/`, run the pipeline scripts
in order. Each script is numbered for reproducibility:

```{.bash}
Rscript scripts/000-staging-check.R         # confirm source data present
Rscript scripts/010-ingest-nice.R           # parse NICE -> JSON-LD
Rscript scripts/010-ingest-dcwf.R           # ... and so on per framework
Rscript scripts/020-assemble-jsonld.R       # uniform parser pass
Rscript scripts/025-export-ntriples.R       # combined RDF graph
Rscript scripts/030-load-graph.R            # smoke-test SPARQL load
```

The combined graph lands at `data/processed/ntriples/_combined.nt`.
Every query on this site reads from that file.

## First query

```{r}
#| label: first-query
#| echo: true
library(cybedtools)
library(dplyr)

# Per-framework summary, lazy-loaded with the package
framework_summary |>
  select(
    framework_name,
    organizing_unit_count,
    elements_per_organizing_unit_with_examples
  ) |>
  arrange(desc(elements_per_organizing_unit_with_examples))
```

## Next steps

- Read [the concepts](../concepts/vocabulary.qmd) page for the
  two-tier schema architecture.
- Browse [the framework pages](../frameworks/index.qmd) for
  per-framework provenance and structure.
- Step through [the canonical queries](../queries/index.qmd) for
  worked examples.
