---
pagetitle: "Concordance · many frameworks, one queryable graph"
toc: false
page-layout: article
---

::: {.hero}

::: {.hero-mark}
![](assets/branding/concordance-lockup-stacked-paper.svg){fig-alt="Concordance: many frameworks, one graph" width=560 .hero-logo}
:::

# Many cybersecurity frameworks. One queryable graph. {.visually-hidden}

```{r}
#| label: setup
#| include: false
suppressPackageStartupMessages(library(dplyr))

fs <- cybedtools::framework_summary |>
  arrange(desc(elements_per_organizing_unit_with_examples))

# Per-framework presentation facts the data does not carry: the page each
# framework links to, and the short name used in the chart and the cards.
# The three v0.3.0 frameworks have no page of their own yet, so they link
# to their section of the package's data-sources documentation.
docs_url <- "https://github.com/ryanstraight/cybedtools/blob/main/docs/framework-data-sources.md"
page <- c(
  "nice-v2"           = "frameworks/nice-v2.html",
  "dcwf-v5.1"         = "frameworks/dcwf-v51.html",
  "ecsf-v1"           = "frameworks/ecsf-v1.html",
  "sfia-9"            = "frameworks/sfia-9.html",
  "cyberorg-k12-v1.0" = "frameworks/cyberorg-k12.html",
  "csta-2017"         = "frameworks/csta-2017.html",
  "csec2017-v1"       = "frameworks/csec2017.html",
  "digcomp-2.2"       = "frameworks/digcomp-22.html",
  "cyqual-v1.2.0"     = paste0(docs_url, "#cyqual-czech-republic-national-cybersecurity-qualifications-framework"),
  "ccssf-2022"        = paste0(docs_url, "#ccssf-canada-canadian-cyber-security-skills-framework"),
  "otccf-v1.1"        = paste0(docs_url, "#otccf-singapore-operational-technology-cybersecurity-competency-framework")
)
short <- c(
  "nice-v2"           = "NICE Framework v2.2.0",
  "dcwf-v5.1"         = "DCWF v5.1",
  "ecsf-v1"           = "ENISA ECSF",
  "sfia-9"            = "SFIA 9",
  "cyberorg-k12-v1.0" = "Cyber.org K-12",
  "csta-2017"         = "CSTA K-12 CS",
  "csec2017-v1"       = "ACM/IEEE CSEC2017",
  "digcomp-2.2"       = "DigComp 2.2",
  "cyqual-v1.2.0"     = "CyQUAL 1.2.0",
  "ccssf-2022"        = "CCSSF 2022",
  "otccf-v1.1"        = "OTCCF v1.1"
)
# Editorial flag, not a property of the data. It marks the one bar whose
# strict-to-with-examples jump the chart annotates as an outlier.
outlier <- c("csta-2017")

juris_label <- c(US = "US", EU = "EU", global = "Global",
                 CZ = "CZ", CA = "CA", SG = "SG")
juris_icon <- c(US = "icon-us", EU = "icon-eu", global = "icon-global",
                CZ = "icon-cz", CA = "icon-ca", SG = "icon-sg")

fs <- fs |>
  mutate(
    href      = unname(page[framework_slug]),
    label     = unname(short[framework_slug]),
    j_label   = unname(juris_label[jurisdiction]),
    j_icon    = unname(juris_icon[jurisdiction]),
    t_label   = tools::toTitleCase(framework_type),
    t_icon    = paste0("icon-", framework_type),
    is_outlier = framework_slug %in% outlier
  )

denom <- max(fs$elements_per_organizing_unit_with_examples)
denom_name <- fs$label[which.max(fs$elements_per_organizing_unit_with_examples)]
lede_names <- paste(fs$label[order(match(fs$framework_slug, names(short)))], collapse = ", ")
```

::: {.lede}
Compare structure, density, and coverage across `r lede_names`. The differences each
framework was authored to express stay intact.
:::

::: {.cta-row}
[Why this exists](why.qmd){.btn .btn-primary}
[Browse frameworks](frameworks/index.qmd){.btn .btn-outline-secondary}
:::

:::

## The same question, different answers

::: {.chart-subtitle}
Granularity per organizing unit, by design. The spread runs roughly
`r round(denom / min(fs$elements_per_organizing_unit_with_examples))` to one across the corpus. It tracks what each framework was
authored to specify, not relative quality.
:::

```{r}
#| label: fig-density
#| results: asis
cat('<div class="density-legend">\n',
    '  <span class="legend-item"><span class="swatch swatch-workforce"></span>Workforce</span>\n',
    '  <span class="legend-item"><span class="swatch swatch-pedagogy"></span>Pedagogy</span>\n',
    '</div>\n', sep = "")
cat('<figure class="density-figure">\n  <div class="density-chart">\n')
for (i in seq_len(nrow(fs))) {
  r <- fs[i, ]
  v <- r$elements_per_organizing_unit_with_examples
  cat(sprintf('    <div class="bar-row" data-type="%s"%s>\n',
              r$framework_type, if (r$is_outlier) " data-outlier" else ""))
  cat('      <div class="bar-meta">\n        <span class="bar-pictograms">\n')
  cat(sprintf('          <svg class="bar-icon" role="img" aria-label="%s"><use href="#%s"/></svg>\n',
              r$j_label, r$j_icon))
  cat(sprintf('          <svg class="bar-icon" role="img" aria-label="%s"><use href="#%s"/></svg>\n',
              r$t_label, r$t_icon))
  cat('        </span>\n')
  cat(sprintf('        <a class="bar-name" href="%s">%s</a>\n', r$href, r$label))
  cat('      </div>\n      <div class="bar-visual">\n')
  cat(sprintf('        <div class="bar-fill" style="width: %s%%;"></div>\n',
              format(round(v / denom * 100, 1), nsmall = 1, trim = TRUE)))
  cat(sprintf('        <span class="bar-value">%s</span>\n',
              format(round(v, 1), nsmall = 1, trim = TRUE)))
  cat('      </div>\n    </div>\n')
}
cat('  </div>\n')
cat(sprintf('  <figcaption>Elements per top-level organizing unit, with examples included. Bar widths are scaled to the highest value (%s, %s).</figcaption>\n',
            denom_name, format(round(denom, 1), nsmall = 1, trim = TRUE)))
cat('</figure>\n')
```

::: {.column-margin}
NICE specifies tasks, knowledge statements, and skill statements per
work role. DigComp specifies competence areas designed for citizen
self-assessment. Different questions, different shapes.
:::

[See the full cross-framework analysis](queries/density-spread.qmd){.no-icon}.

## The frameworks at a glance

```{r}
#| label: framework-cards
#| results: asis
cat('<div class="framework-grid">\n')
for (i in seq_len(nrow(fs))) {
  r <- fs[i, ]
  cat(sprintf('  <a class="framework-card" href="%s">\n', r$href))
  cat('    <span class="pictograms">\n')
  for (p in list(c(r$j_icon, r$j_label), c(r$t_icon, r$t_label))) {
    cat('      <span class="pictogram">\n')
    cat(sprintf('        <span class="pictogram-circle"><svg class="pictogram-icon" aria-hidden="true"><use href="#%s"/></svg></span>\n', p[1]))
    cat(sprintf('        <span class="pictogram-label">%s</span>\n', p[2]))
    cat('      </span>\n')
  }
  cat('    </span>\n')
  cat(sprintf('    <span class="name">%s</span>\n', r$label))
  cat(sprintf('    <span class="stats"><span class="label">Units</span> %s \u00b7 <span class="label">Elements</span> %s</span>\n',
              format(r$organizing_unit_count, big.mark = ",", trim = TRUE),
              format(r$element_count_with_examples, big.mark = ",", trim = TRUE)))
  cat('  </a>\n')
}
cat('</div>\n')
```

## What this is and isn't

::: {.callout-defensive role="note" aria-label="What this is and isn't"}

`cybedtools` is a comparison and querying layer over existing
cybersecurity workforce and learning frameworks. **When the package's
outputs disagree with a framework steward's own view of their framework,
the steward's view is correct.**

The package extends framework-based analysis. It gives researchers a
uniform query surface across heterogeneous artefacts. It does not
replace any framework, redistribute source text under more permissive
terms than the framework's own license, or rank frameworks against
each other.

:::
