---
title: "Using cybedtools with an AI assistant"
description: >
  Where the machine-readable documentation lives, and the rules an assistant
  has to follow to write cybedtools code that runs.
---

```{r}
#| include: false
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This article has two readers: a person who wants an assistant to write cybedtools code, and the assistant itself, reading its Markdown twin. An assistant working from general R and SPARQL knowledge will get several things about this package wrong.

`AGENTS.md` in the repository root carries the same rules for coding agents working inside a clone.

## Machine-readable documentation

The site publishes a reference index at <https://ryanstraight.github.io/cybedtools/llms.txt>, and every reference page and article has a Markdown twin at the same URL with `.md` appended. The index points at pages, not at the rules below.

The documentation is also indexed in Context7 at <https://context7.com/ryanstraight/cybedtools>. If your assistant supports Context7, point it there before it writes code.

## Rules

### Use the domain helpers, and keep SPARQL to one triple pattern

```{r}
#| eval: false

library(cybedtools)
library(dplyr)

role_framework_bindings(rdf) |>
  inner_join(role_element_bindings(rdf), by = "role") |>
  count(framework_name, name = "element_count")
```

The `librdf` C library that `rdflib` wraps hangs for many minutes on multi-pattern SPARQL joins and silently returns zero rows on multi-property selects against a single subject. Single basic graph patterns execute fast and correctly. `sparql_pairs()` and `sparql_subjects()` issue those single patterns; the domain helpers compose them and join in dplyr. Never write `COUNT`, `GROUP BY`, or `HAVING` into a SPARQL string. Aggregate in dplyr.

### Load the graph with a loader, and know the package does not ship one

```{r}
#| eval: false

rdf <- make_demo_graph()

rdf_full <- load_combined_ntriples_graph()
```

The installed package does not bundle the framework graph. It ships ingestion code and no framework text. `make_demo_graph()` returns a synthetic two-framework graph that exercises every domain helper without staged data. `load_combined_ntriples_graph()` needs a graph you assembled from staged sources.

### Join on IRIs, never on bare statement codes

```{r}
#| eval: false

element_framework_bindings(rdf_full) |>
  inner_join(element_text(rdf_full), by = "element")
```

Statement codes are unique only within a framework. NICE and CCSSF both print codes in the `T0516` shape, and they denote different statements. CyQUAL reuses the 2017 NICE task codes and shares the `T1xxx` range with current NICE for unrelated text. Join on the full IRI, or carry a framework column alongside the code.

`element_text()` returns `element` and `text`. Parsed sub-points and examples are elements and carry text, so parent-statement analyses must anti-join `subpoint_framework_bindings()` and `example_framework_bindings()`, renaming their `subpoint` and `example` columns to `element`.

### Restrict to roles before counting per role

```{r}
#| eval: false

role_element_bindings(rdf_full) |>
  semi_join(role_framework_bindings(rdf_full), by = "role") |>
  count(role, name = "element_count")
```

`role_element_bindings()` returns bindings for every organizing unit, not only roles, because `cybed:hasElement` is the universal parent-child link. A role-level count made without that `semi_join` silently includes SFIA skills and DigComp competence areas.

### Say which element count you mean

```{r}
#| eval: false

cybedtools::framework_summary |>
  select(framework_name, element_count_strict, element_count_with_examples)
```

`element_count_with_examples` counts parents plus `cybed:Subpoint` children plus `cybed:Example` children. `element_count_strict` counts parents only. The README headline density finding uses the with-examples count. The related count in `docs/framework-invariants.yml`, `total_elements_with_subpoints`, counts parents plus Subpoints and excludes Examples.

### Pick the right cross-framework type

```{r}
#| eval: false

organizing_unit_framework_bindings(rdf_full)

role_framework_bindings(rdf_full)
```

`cybed:OrganizingUnit` reaches all eleven frameworks. `cybed:Role` reaches the six that declare roles: NICE, DCWF, ECSF, CyQUAL, CCSSF, and OTCCF. A query written against `cybed:Role` when the question is corpus-wide drops five frameworks without saying so.

### Respect the upstream licenses

```{r}
#| eval: false

framework_metadata(rdf_full) |>
  select(name, jurisdiction, sector)
```

Package code is MIT. Framework content keeps its upstream terms. `docs/framework-invariants.yml` records a `public_redistribution` policy per framework: SFIA is `local_only`, CSEC2017 and OTCCF are `structure_only`, CyQUAL and CCSSF are `full_with_attribution`. OTCCF, CyQUAL and CCSSF are in the package by the written permission of their stewards. CCSSF's attribution is "Copyright Government of Canada. Used with permission of the Canadian Centre for Cyber Security." Do not reproduce statement text from a framework marked `local_only` or `structure_only`. Titles, counts, categories and mappings are fine. See `LICENSE.md` and `LICENSING.md`.

### Cite the package and the stewards

Cite the concept DOI, `10.5281/zenodo.20076116`, which resolves to the latest release. Cite the framework stewards for framework content: CSA for OTCCF, CyQUAL and Masaryk University for CyQUAL, and the Canadian Centre for Cyber Security for CCSSF, which asks that its material be referenced when used. Do not credit the assistant.

## A preamble you can paste

> When writing cybedtools code: use the domain helpers (`framework_metadata`, `organizing_unit_framework_bindings`, `role_framework_bindings`, `element_framework_bindings`, `role_element_bindings`, `subpoint_framework_bindings`, `example_framework_bindings`, `element_text`) and the primitives `sparql_pairs` and `sparql_subjects`. Any raw `rdflib::rdf_query` must be a single basic graph pattern, joined in dplyr; librdf hangs on multi-pattern joins and silently returns zero rows on multi-property selects. Load with `load_combined_ntriples_graph()`, or `make_demo_graph()` when no data is staged; the installed package bundles no framework graph. Never join frameworks on a bare element code, only on full IRIs or framework-qualified ids. `role_element_bindings()` covers every organizing unit, so restrict it with a `semi_join` on `role_framework_bindings()` before counting per role. State whether a count is strict or with-examples. Do not reproduce framework statement text. Cite DOI 10.5281/zenodo.20076116 and the framework stewards.

## What not to ask for

Do not ask an assistant to produce a crosswalk or alignment the graph does not contain. CCSSF's citations of 2017-era NICE work roles are carried as `cybed:niceCrossReference` literals, never as links, and they do not match the NICE v2.2.0 identifiers in this graph. An assistant that invents the missing edges is fabricating.

Do not treat a similarity score as equivalence. The Concordance alignment tables report full-document Jaccard vocabulary overlap. A best match surfaces a structural candidate. Credentialing pathways, regulatory contexts, clearance requirements and language fluency are separate questions, and the equivalence judgment is human work.
